---
title: "Customize Specification Curve Plot"
output: github_document
---

This vignette exemplifies two ways to plot the specification curve.


## Run the specification curve analysis

In order to have some data to work with, we run the minimal example contained in the package.

```{r, message = F, warnings = F}
library(specr)
library(dplyr)
# run spec analysis
results <- run_specs(example_data,
                     y = c("y1", "y2"),
                     x = c("x1", "x2"),
                     model = "lm",
                     controls = c("c1", "c2"),
                     subset = list(group1 = unique(example_data$group1),
                                   group2 = unique(example_data$group2)))
```

Let's quickly get some ideas about the specification curve by using `summarise_specs()`

```{r, message = F, warnings = F}
# Overall median
summarise_specs(results)

# By choices
summarise_specs(results, group = c("x", "controls"))
```

We see that it makes quite a difference whether `x1` or `x2` is used as independent variable. 


## A simple way to visualize the results 

The simplest way to visualize most of the information contained in the results data frame is by using the `plot_specs()` function. 

```{r, fig.height=10, fig.width=10, message=F, warning = F}
plot_specs(results)
```

We can further customize that function, e.g., by removing unnecessary information or by reordering/transforming the analytical choices (and thereby visualize specific contrasts).

```{r, fig.height=10, fig.width=10, message=F, warning = F}
plot_specs(results, 
           choices = c("x", "y", "controls", "subsets"),  # model is not plotted
           rel_heights = c(0.5, 2)) # changing relative heights of the two parts

results %>%
  mutate(group1 = ifelse(grepl("group1 = 0", subsets), "0", "1"),
         group2 = ifelse(grepl("group2 = A", subsets), "A", "B & C")) %>%
  plot_specs(choices = c("x", "y", "controls", "group1", "group2"))
```


## A more customizable way to visualize the results

Alternatively, we can also plot the curve and the choice panel individually and bind them together afterwards.

```{r, fig.height=10, fig.width=10, message=F, warning = F}
# Plot specification curve (p1)
p1 <- plot_curve(results) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey") +
  ylim(-8, 12) +
  labs(x = "", y = "unstandarized regression coefficient") +
  theme_grey()

# Plot choices (p2)
p2 <- plot_choices(results) +
  labs(x = "specifications (ranked)") +
  theme_grey()

# Combine plots
plot_specs(plot_a = p1,
           plot_b = p2,
           labels = c("", ""),
           rel_height = c(2, 2.5))
```

## Include sample size histogram

The additional function `plot_samplesizes()` provides and additional panel that can be added via the `cowplot::plot_grid()`

```{r, fig.height=10, fig.width=10, message=F, warning = F}
p3 <- plot_samplesizes(results) +
  theme_grey()

# Combine via cowplot
cowplot::plot_grid(p1, p2, p3,
                   ncol = 1,
                   align = "v",
                   rel_heights = c(1.5, 2, 0.8),
                   axis = "rbl")
```



