---
title: "Including random effects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Including random effects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  #fig.path = "man/figures/README-",
  out.width = "100%",
  fig.retina = 2
)
```

Many data are hierarchical and we want to acknowledge the nested structure in our models. Specr can easily estimate such multilevel models. We again have to write a customized function that we can pass first to `setup()` and the `specr()` will do the rest. 

For this example, we will us the `gapminder` data set that is included in the `gapminder`-package. We quickly recode some variable to get more interpretable estimates.

```{r, message=F, warning = F}
library(specr)
library(gapminder)
library(tidyverse)

# Recode some variables
gapminder <- gapminder %>%
  mutate(gdpPercap_log = log(gdpPercap),
         pop = pop/1000)

# Check data
gapminder
```

## 1. Defining a customatized multilevel function

For this example, we use the package `lme4` and more specifically the function `lmer()` to estimate the multilevel model (more complex models such as poisson or negative binomial multilevel models can likewise be estimated). 

Based on the data set, we want to estimate the relationship between `gdpPercap` (GDP per capita) and `lifeExp` (life expectancy). Both variables are nested within both countries and years. We create several customized lmer-functions that account for different nesting strucrures. 

Several aspects are important to take into account: 
- You need to load `lavaan` within the function to be able to use `lmer()`
- You need to load `broom.mixed` as the standard `broom::tidy` function does not work for merMod-objects.


```{r}
# Random intercept model (only country as grouping variable)
lmer_ri_1 <- function(formula, data,...) {
  require(lme4)
  require(broom.mixed)
  formula <- paste(formula, "+ (1|country)")
  lmer(formula, data)
}

# Including random slopes (only country as grouping variable)
lmer_rs_1 <- function(formula, data,...) {
  require(lme4)
  require(broom.mixed)
  slopevars <- unlist(strsplit(formula, " ~ "))[2]
  formula <- paste0(formula, "+ (1 + ", slopevars, "|country)" )
  lmer(formula, data)
}

# Random intercept model (lifeExp is nested in both countries and years)
lmer_ri_2 <- function(formula, data,...) {
  require(lme4)
  require(broom.mixed)
  formula <- paste0(formula, "+ (1|country) + (1|year)")
  lmer(formula, data)
}

# Including random slopes (intercept and slopes are nested in both countries and years)
lmer_rs_2 <- function(formula, data,...) {
  require(lme4)
  require(broom.mixed)
  slopevars <- unlist(strsplit(formula, " ~ "))[2]
  formula <- paste0(formula, "+ (1 + ", slopevars, "|country) + (", slopevars, "|year)" )
  lmer(formula, data)
}
```

## 2. Setup the models

We can now use these function to estimate these models. In this example, we investigate the influence of differnet nesting structure on the fixed effect between GDP per capita and life expectancy.

```{r, message=F, warning = F}
# Setup specifications with customized functions
specs <- setup(data = gapminder,
                y = c("lifeExp"),
                x = c("gdpPercap_log"), 
                model = c("lmer_ri_1", "lmer_ri_2", 
                          "lmer_rs_1", "lmer_rs_2"))

# Check specifications
summary(specs, print.specs = T)
```

# 3. Fit the models

```{r}
opts <- furrr_options(
  globals = list(lmer_ri_1 = lmer_ri_1, lmer_ri_2 = lmer_ri_2, lmer_rs_1 = lmer_rs_1, lmer_rs_2 = lmer_rs_2),
  packages = c("lme4", "broom.mixed")
)

results <- specr(specs, workers = 2)
summary(results)
```

And we can again plot our results.

```{r, fig.height=8, fig.width=8, message=F, warning = F}
# Plot
plot(results, choices = c("x", "y", "model"))
```

